Human Factors Influencing Visual Statistical Inference

### 2013-06-06 sociology_paper.tex
model formula

@@

As per the design of the experiment for learning trend we have two different response variables. One is observer feedback on a lineup which is a binary variable.  The other variable is the time taken for each evaluation which is a continuous variable. Suppose random variable $Y_{\ell, i}$ denotes the response from subject $ i=1,2, ..., K$ on lineup $\ell=1,2, ..., L$. We model the mean $\mu_{\ell, i} = E(Y_{\ell, i})$ with a generalized liner mixed effect model

@@

\noindent where $g(.)$ denotes the link function. $W$ is a design matrix of covariates corresponding to specifics of lineup $\ell$ and subject $i$, and $\delta$ is the vector of corresponding parameters. Covariates could include  demographic information of individuals, such as age, gender, education level etc.,  as well as the attempt made by the observer. $Z_{\ell i}$,  $1 \le i \le K$, $1 \le \ell \le L$,  is a design matrix corresponding to random effects specific to individual $i$ and lineup $\ell$; and 
$\tau$  is a vector of independent normally distributed random variables $\tau_{\ell i}$ with  variance matrix $\sigma_\tau I_{KL \times KL}$. $\tau$ will usually include a component incorporating an individual's ability or skill to evaluate lineups.

>>>>

\usepackage{color}
\newcommand{\hh}[1]{{\color{magenta} #1}}

@@

As per the design of the experiment for learning trend we have two different response variables. One is observer feedback on a lineup which is a binary variable.  The other variable is the time taken for each evaluation which is a continuous variable. Suppose random variable $Y_{\ell, i}$ denotes the response from subject $ i=1,2, ..., I$ on lineup $\ell=1,2, ..., L$. We model the mean $\mu_{\ell, i} = E(Y_{\ell, i})$ with a generalized liner mixed effect model

@@

\hh{
Let $\pi_{ijk}$ be the probability that  observer $i$ picks the data panel from lineup $j$ in his/her $k$th attempt. We model this in a mixed effects model of the form
\begin{equation} \label{eqn:trend}
g( \pi_{ijk} )= \mu + \alpha_k + u_i +  a_{i} k + \ell_j,  
\end{equation}
where $\mu$ is an overall average probability for picking out the data plot from a lineup, $\alpha_k$ is the effect of the $k$th attempt on the probability, with $\alpha_1 = 0$ and $k = 1, ..., K$.
$u_i$ and $a_i$ are observer specific random effects, $i = 1, ..., I$. $u_i$ is a random intercept, describing a basic subject-specific ability. We assume $u_i \sim N(0, \sigma_u^2)$. 
$a_i$ is a random slope capturing an individuals' specific learning effect over the course of $K$ attempts. We assume $a_i \sim N(0, \sigma_a^2)$. 
For $\ell_j$ we again assume a normal distribution, $N(0, \sigma_\ell^2)$. $\ell_j$ is a random intercept predicting lineup difficulty level.
}

\noindent  $g(.)$ denotes the link function. $W$ is a design matrix of covariates corresponding to specifics of lineup $\ell$ and subject $i$, and $\delta$ is the vector of corresponding parameters. Covariates could include  demographic information of individuals, such as age, gender, education level etc.,  as well as the attempt made by the observer. $Z_{\ell i}$,  $1 \le i \le K$, $1 \le \ell \le L$,  is a design matrix corresponding to random effects specific to individual $i$ and lineup $\ell$; and 
$\tau$  is a vector of independent normally distributed random variables $\tau_{\ell i}$ with  variance matrix $\sigma_\tau I_{KL \times KL}$. $\tau$ will usually include a component incorporating an individual's ability or skill to evaluate lineups.
